A Simple Type Theory for Category Theory

### 2022-05-25 paper/paper.tex
work on related work

  \begin{enumerate}
  \item The indexing cateogry is $\Set$, which is a category with families.
  \item The equipment is defined as $\mathcal U\Gamma = \mathcal V^\Gamma$.
  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
    given pointwise.
  \item The representing object for objects is simply
    $\textrm{Ob}_\Gamma = \textrm{Ob}(\mathcal V)^\Gamma$ and
    similarly for the other universe objects.
  \end{enumerate}
\end{definition}

@@

\paragraph{Calculi for Presheaves/Profunctors}

David Jaz Myers String Diagrams

Coend calculus

Cartesian Bicategories.

@@

\paragraph{Directed Type Theory}

@@

\paragraph{Category Theoretic Extensions}

@@

\paragraph{Semantics}

>>>>

  \begin{enumerate}
  \item The indexing cateogry is $\Set$, which is a category with families.
  \item The equipment is defined as $\mathcal U\Gamma = \mathcal V^\Gamma$.
  \item Restriction by a function $\gamma : \Delta \to \Gamma$ is
    given pointwise.
  \item The representing object for objects is $\textrm{Ob}_\Gamma =
    \textrm{Ob}(\mathcal V)^\Gamma$ and similarly for the other
    universe objects.
  \end{enumerate}
\end{definition}

@@

\paragraph{Calculi for 2-dimensional category theory}

\citet{djm-string-diagrams} provides a string diagram calculus for
double categories and pro-arrow equipments, generalizing string
diagrams for monoidal categories \citet{joyal?}. These are an
alternative approach to type theoretic calculi, with the string
diagrams typically making tensor products simpler to work with, while
a type theoretic calculus like SCT makes the closed structure
$\homrallXYtoZ \alpha P Q$ simpler to work with by using bound
variables.

One alternative to proarrow equipments is to axiomatize the bicategory
of categories and profunctors rather than the double category of
categories functors and profunctors. One abstract structure along
these lines is a \emph{Cartesian bicategory}\footnote{which is not the
same as a bicategory with products/finite limits, for which a syntax
was developed \citet{fioresaville}} is given in
\citet{carboni-walters,carboni-walters-someonelese}. Some preliminary
work on a proof system for Cartesian bicateogires was presented at
\citet{jonasfreycartesianbicategories}. This bears strong similarities
to our calculus SCT, except that (1) their profunctors may have 0, 1
or more covariant or contravariant variables but (2) they do not have
a term syntax like our natural transformation syntax and (3) they do
not have functors and restriction of profunctors by functors (since
this structure is not present in a Cartesian
bicategory). Semantically, Cartesian bicategories are somewhat less
general than pro-arrow equipments because in general the set of
functors between two categories can only be recovered from the set of
profunctors if the codomain category is Cauchy complete, in which case
functors are equivalent to adjoint pairs of profunctors.

Our type-theoretic syntax for virtual double categories is intended to
be a formal notation for the manipulation of profunctors. As such it
has similarities to informal syntax for manipulating profunctors
involving ends and coends, such as the coend calculus of
\citet{coend-calculus}.

Finally, some variations on double categories have been used to model
the structure of certain program logics. \citet{newlicata20} present a
logic for \emph{vertically thin} pro-arrow equipments, i.e., pro-arrow
equipments where the functor and natural transformation judgments are
posetal. This calculus is quite different from ours in that the
horizontal morphisms are terms of a programming language, and so they
are not presented as having co-variant and contravariant variables,
and the equipment structure is modeled by explicit coercions rather
than by syntactic substitution of vertical arrows in horizontal
arrows. Another similar calculus is System P \citep{systemP} which is
an internal language of \emph{reflexive graph categories}, which are
like double categories without horizontal composition.

@@

\subsection{Directed Type Theory}

@@

\subsection{Category Theoretic Extensions}

@@

\subsection{Semantics}
